\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-07-01 paper.qmd
Added best fit projections

The appropriate model for the S-curve using UMAP (n_neighbors: $15$) was created with $420$ bins. It was determined to be the best model after removing edges with a length greater than $0.134$. The decision was made considering that the model accurately captures the geometry of the S-curve. But better in some places only (@fig-scurve_sc_best).

In contrast, a poor model for the S-curve with PHATE was created using only $190$ bins and removing edges with a length greater than $0.3$. This model is inadequate as it has a higher MSE compared to the appropriate model. Visually, the model squeezed to the middle of the geometry of the S-curve (@fig-scurve-sc-bad).

XXX BEST FIT FOR S-CURVE MIGHT BE A DIFFERENT NLDR. 
SHOW THE SEVERAL SUB-OPTIMAL FITS.

>>>>

@fig-scurve-sc-best shows one of the best fits for S-curve data. TriMAP with $15$ nearest neighbors for forming the nearest neighbor triplet (`n_inliers`), $3$ outliers for forming the nearest neighbor triplets (`n_outliers`), $3$ random triplets per point (`n_random`).

<!--langevitour with pD model for S-curve-->
::: {#fig-scurve-sc-best layout-ncol="4" fig-pos="H"}
![](figures/scurve/sc_best_trimap_trimesh.png)

![](figures/scurve/sc_trimap_best_1.png)

![](figures/scurve/sc_trimap_best_2.png)

![](figures/scurve/sc_trimap_best_3.png)

Model in \gD{}, on the TriMAP layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data. The model closely fits the shape, and it adequately captures the surface. (The **langevitour** software is used to view the data with a tour, and the full video is available at (<https://youtu.be/zEz2MUCtADM>). 
:::

<!-- The appropriate model for the S-curve using UMAP (n_neighbors: $15$) was created with $420$ bins. It was determined to be the best model after removing edges with a length greater than $0.134$. The decision was made considering that the model accurately captures the geometry of the S-curve. But better in some places only (@fig-scurve_sc_best). -->

<!-- In contrast, a poor model for the S-curve with PHATE was created using only $190$ bins and removing edges with a length greater than $0.3$. This model is inadequate as it has a higher MSE compared to the appropriate model. Visually, the model squeezed to the middle of the geometry of the S-curve (@fig-scurve-sc-bad). -->

<!-- XXX BEST FIT FOR S-CURVE MIGHT BE A DIFFERENT NLDR.  -->
<!-- SHOW THE SEVERAL SUB-OPTIMAL FITS. -->

### 2024-07-01 paper.tex
Added best fit projections

The appropriate model for the S-curve using UMAP (n\_neighbors: \(15\))
was created with \(420\) bins. It was determined to be the best model
after removing edges with a length greater than \(0.134\). The decision
was made considering that the model accurately captures the geometry of
the S-curve. But better in some places only
(\textbf{?@fig-scurve\_sc\_best}).

In contrast, a poor model for the S-curve with PHATE was created using
only \(190\) bins and removing edges with a length greater than \(0.3\).
This model is inadequate as it has a higher MSE compared to the
appropriate model. Visually, the model squeezed to the middle of the
geometry of the S-curve (\textbf{?@fig-scurve-sc-bad}).

XXX BEST FIT FOR S-CURVE MIGHT BE A DIFFERENT NLDR. SHOW THE SEVERAL
SUB-OPTIMAL FITS.

>>>>

Figure~\ref{fig-scurve-sc-best} shows one of the best fits for S-curve
data. TriMAP with \(15\) nearest neighbors for forming the nearest
neighbor triplet (\texttt{n\_inliers}), \(3\) outliers for forming the
nearest neighbor triplets (\texttt{n\_outliers}), \(3\) random triplets
per point (\texttt{n\_random}).

\begin{minipage}{0.25\linewidth}
\includegraphics{figures/scurve/sc_best_trimap_trimesh.png}\end{minipage}%
%
\begin{minipage}{0.25\linewidth}
\includegraphics{figures/scurve/sc_trimap_best_1.png}\end{minipage}%
%
\begin{minipage}{0.25\linewidth}
\includegraphics{figures/scurve/sc_trimap_best_2.png}\end{minipage}%
%
\begin{minipage}{0.25\linewidth}
\includegraphics{figures/scurve/sc_trimap_best_3.png}\end{minipage}%

\caption{\label{fig-scurve-sc-best}Model in \gD{}, on the TriMAP layout,
and three views of the fit in projections from \(7\text{-}D\), for the
S-curve data. The model closely fits the shape, and it adequately
captures the surface. (The \textbf{langevitour} software is used to view
the data with a tour, and the full video is available at
(\url{https://youtu.be/zEz2MUCtADM}).}
